\documentclass[sigconf]{acmart}

\AtBeginDocument{ \providecommand\BibTeX{ Bib\TeX } }
\setcopyright{acmlicensed}
\copyrightyear{2025}
\acmYear{2025}
\acmDOI{XXXXXXX.XXXXXXX}

\acmConference[BI 2025]{Business Intelligence}{-}{-}

\begin{document}

\title{BI2025 Experiment Report - Group 43}
%% ---Authors: Dynamically added ---

          \author{Johanna Six}
          \authornote{Student A, Matr.Nr.: 12019873}
          \affiliation{
            \institution{TU Wien}
            \country{Austria}
          }
          
          \author{Bernhard Siegl}
          \authornote{Student B, Matr.Nr.: 76543210}
          \affiliation{
            \institution{TU Wien}
            \country{Austria}
          }
          
          \author{Bernhard Siegl}
          \authornote{Student B, Matr.Nr.: 12120509}
          \affiliation{
            \institution{TU Wien}
            \country{Austria}
          }
          

\begin{abstract}
  This report documents the machine learning experiment for Group 43, following the CRISP-DM process model.
\end{abstract}

\ccsdesc[500]{Computing methodologies~Machine learning}
\keywords{CRISP-DM, Provenance, Knowledge Graph, Machine Learning}

\maketitle

%% --- 1. Business Understanding ---
\section{Business Understanding}

\subsection{Data Source and Scenario}
The data source consists of 21 attributes and about 3000 rows. It contains data from December 2019 and 2020, that consists of the number of delayed, cancelled or diverted flights and the cause of delays. The data can be used by airlines to identify what causes the most flight delays / cancellations.

\subsection{Business Objectives}
The airline wants to improve punctuality and therefore reduce the number of delayed, cancelled flights.

%% --- 2. Data Understanding ---
\section{Data Understanding}
\textbf{Dataset Description:} Airline delay counts per carrier per airport for December 2019 and December 2020.

The following features were identified in the dataset:

\begin{table}[h]
  \caption{Raw Data Features}
  \label{tab:features}
  \begin{tabular}{lp{0.2\linewidth}p{0.4\linewidth}}
    \toprule
    \textbf{Feature Name} & \textbf{Data Type} & \textbf{Description} \\
    \midrule
    airport & string> & Three-letter airport code for arrival airport. \\
    airport\_name & string> & Full name of the airport. \\
    arr\_cancelled & integer> & Number of cancelled flights. \\
    arr\_del15 & integer> & Flights arriving more than 15 minutes late. \\
    arr\_delay & double> & Total delay time in minutes. \\
    arr\_diverted & integer> & Number of diverted flights. \\
    arr\_flights & integer> & Number of flights arriving at airport. \\
    carrier & string> & Two-letter airline carrier code. \\
    carrier\_ct & integer> & Flights delayed due to air carrier (e.g. no crew). \\
    carrier\_delay & double> & Delay minutes due to air carrier issues. \\
    carrier\_name & string> & Full airline carrier name. \\
    late\_aircraft\_ct & integer> & Flights delayed due to a previous late aircraft. \\
    late\_aircraft\_delay & double> & Delay minutes due to previous late aircraft. \\
    month & integer> & Month (1–12) of data collection. \\
    nas\_ct & integer> & Flights delayed due to National Aviation System. \\
    nas\_delay & double> & Delay minutes due to National Aviation System. \\
    security\_ct & integer> & Flights canceled due to security issues. \\
    security\_delay & double> & Delay minutes due to security issues. \\
    weather\_ct & integer> & Flights delayed due to weather. \\
    weather\_delay & double> & Delay minutes due to weather. \\
    year & gYear> & Year in which flight data was collected. \\
    \bottomrule
  \end{tabular}
\end{table}

%% --- 3. Data Preparation ---
\section{Data Preparation}
\subsection{Handling outliers}
Data Preparation

\subsection{Considered but Not Applied Pre-processing Steps}
Several preprocessing steps were considered but deliberately not applied:

- Outlier removal: The numerical variables are extreme right-skewed, so those values correspond to large airports, carriers or extreme delay events. If we would remove those, we would distort the business objective and underrepresent the major datapoints. We want to reserve the real world variability.

- Scaling/Log transformations: We decided to not scale some columns yet, since me may do that in the modeling phase if it is required by an algorithm

- Removal of attributes due to correlation: We have a strong correlation between arr\_delay and carrier\_delay, nas\_delay, .. and also between  arr\_del15 and other delay counts. We believe the different delay values encode different operational causes so we would remove some information which could be necessary for our business objective.

- Encoding of categorical attributes: We will do that in the modeling phase since it depends on which algorithm we use.
\subsection{Analysis of Potential Derived Attributes}
Several preprocessing steps were considered but deliberately not applied:

- Outlier removal: The numerical variables are extreme right-skewed, so those values correspond to large airports, carriers or extreme delay events. If we would remove those, we would distort the business objective and underrepresent the major datapoints. We want to reserve the real world variability.

- Scaling/Log transformations: We decided to not scale some columns yet, since me may do that in the modeling phase if it is required by an algorithm

- Removal of attributes due to correlation: We have a strong correlation between arr\_delay and carrier\_delay, nas\_delay, .. and also between  arr\_del15 and other delay counts. We believe the different delay values encode different operational causes so we would remove some information which could be necessary for our business objective.

- Encoding of categorical attributes: We will do that in the modeling phase since it depends on which algorithm we use.
\subsection{Analysis of Potential External Data Sources}
Several external data sources were identified as potentially beneficial.

- Weather data: Observations linked to airport and date
- Airport infrastructure data: Characteristics like number of runways, passenger volume, runway utilization
- Airline operational characteristics: fleet size and age, network complexity
- Public holiday and event calendar: Flag major travel periods or high demand spikes caused by events
- Air traffic control: Information on airspace restrictions or other temporary regulations
\subsection{Outcome of the Data Preparation Phase}
Final dataset after data preparation. Contains only valid, non-negative, and semantically consistent airline delay records. Used as input for the modeling phase.


%% --- 4. Modeling ---
\section{Modeling}

\subsection{Hyperparameter Configuration}
The model was trained using the following hyperparameter settings:

\begin{table}[h]
  \caption{Hyperparameter Settings}
  \label{tab:hyperparams}
  \begin{tabular}{lp{0.4\linewidth}l}
    \toprule
    \textbf{Parameter} & \textbf{Description} & \textbf{Value} \\
    \midrule
    Learning Rate & ... & 1.23 \\
    \bottomrule
  \end{tabular}
\end{table}

\subsection{Training Run}
A training run was executed with the following characteristics:
\begin{itemize}
    \item \textbf{Algorithm:} Random Forest Algorithm
    \item \textbf{Start Time:} 2025-12-15 16:55:45
    \item \textbf{End Time:} 2025-12-15 16:55:45
    \item \textbf{Result:} R-squared Score = 1.2300
\end{itemize}

%% --- 5. Evaluation ---
\section{Evaluation}

%% --- 6. Deployment ---
\section{Deployment}

\section{Conclusion}

\end{document}
